\Large\bf Scalable and Probabilistic Leaderless BFT Consensus through Metastability

### 2020-11-07 avalanche_fr.tex
S5 - début DAG, création de avalanche_fr.tex

Avalanche consists of multiple single-decree Snowball instances instantiated as a multi-decree protocol that
maintains a dynamic, append-only directed acyclic graph (DAG) of all known transactions.
The DAG has a single sink that is the \emph{genesis vertex}.
Maintaining a DAG provides two significant benefits.
First, it improves efficiency, because a single vote on a DAG vertex implicitly votes for all transactions on the path to the genesis vertex.
Second, it also improves security, because the DAG intertwines the fate of transactions, similar to the Bitcoin blockchain.
This renders past decisions difficult to undo without the approval of correct nodes.

% Avalanche's DAG embodies all the transactions that have been proposed to the system, where each transaction is represented as a vertex.
% The DAG is rooted at a protocol-defined, well-known \emph{genesis} vertex.
When a client creates a transaction, it names one or more \emph{parents}, which are included inseparably in the transaction and form the edges of the DAG\@.
The parent-child relationships encoded in the DAG may, but do not need to, correspond to application-specific dependencies; for instance, a child transaction need not spend or have any relationship with the funds received in the parent transaction.
% These additional relationships entangle the fate of previous decisions made by the system.
We use the term \emph{ancestor set} to refer to all transactions reachable via parent edges back in history, and \emph{progeny} to refer to all children transactions and their offspring.

The central challenge in the maintenance of the DAG is to choose among \emph{conflicting transactions}.
The notion of conflict is application-defined and transitive, forming an equivalence relation.
In our cryptocurrency application, transactions that spend the same funds (\emph{double-spends}) conflict, and form a \emph{conflict set}
%(Figure~\tronly{\ref{fig:dag-conflict-set}}{\ref{fig:dag-cd}})
(shaded regions in Figure~\ref{fig:dag-cd}), out of which only a single one can be accepted.
Note that the conflict set of a virtuous transaction is always a singleton.
%Note that the graph structure of transactions that spend depend on each other, also known as the UTXO graph, is completely independent of the DAG that Avalanche maintains. As a result, any two vertices may be in conflict. Figure~\ref{fig:dag-conflicting-set} shows an example.

>>>>

%O Avalanche consists of multiple single-decree Snowball instances instantiated as a multi-decree protocol that
%O maintains a dynamic, append-only directed acyclic graph (DAG) of all known transactions.
Avalanche consiste en plusieurs instance Snowball à décret unique instanciées en tant que protocole multi-décrets qui maintient un graphe orienté acyclique (\emph{Directed Acyclic Graph},  DAG) dynamique en ajout seul de toutes les transactions connues.
%O The DAG has a single sink that is the \emph{genesis vertex}.
Le DAG a un nœud terminal (\emph{sink}) unique qui est le \emph{genesis vertex}, le sommet genèse. % HOULA sink ? 
%O Maintaining a DAG provides two significant benefits.
L'emploi d'un DAG procure deux avantages :
%O First, it improves efficiency, because a single vote on a DAG vertex implicitly votes for all tran
%O sactions on the path to the genesis vertex.
D'abord, il est plus efficient car un vote sur un sommet vote implicitement pour toutes les transactions sur le chemin vers le sommet genèse.
%O Second, it also improves security, because the DAG intertwines the fate of transactions, similar to the Bitcoin blockchain.
Ensuite, il améliore également la sécurité parce que le entrelace le sort des transactions, similaire en cela à la blockchain de Bitcoin.
%O This renders past decisions difficult to undo without the approval of correct nodes.
Cela rend les décisions passées difficiles à défaire sans l'approbation des nœuds corrects.

% Avalanche's DAG embodies all the transactions that have been proposed to the system, where each transaction is represented as a vertex.
% The DAG is rooted at a protocol-defined, well-known \emph{genesis} vertex.
%O When a client creates a transaction, it names one or more \emph{parents}, which are included inseparably in the transaction and form the edges of the DAG\@.
Quand un client crée une transaction, il nomme un ou plusieurs \emph{parents}, et ceux-ci sont inclus de façon inséparable dans la transaction et forment les bords du DAG\@.
%O The parent-child relationships encoded in the DAG may, but do not need to, correspond to application-specific dependencies; for instance, a child transaction need not spend or have any relationship with the funds received in the parent transaction.
Les relations parent-enfant codées dans le DAG peuvent, sans obligation, correspondre à des dépendances spécifiques aux applications ; par exemple, la transaction d'un enfant ne dépense pas nécessairement, ni n'a forcément de relation avec les fonds reçus lors de la transaction parent.
% These additional relationships entangle the fate of previous decisions made by the system.
%O We use the term \emph{ancestor set} to refer to all transactions reachable via parent edges back in history, and \emph{progeny} to refer to all children transactions and their offspring.
Nous utilisons le terme \emph{ensemble ancêtre} pour nous référer à toutes les transactions accessibles par les arêtes parent historisées, et \emph{progéniture} pour nous référer à toutes les transactions enfant et leur descendance.

%O The central challenge in the maintenance of the DAG is to choose among \emph{conflicting transactions}.
Le principal défi de la maintenance du DAG est de choisir parmi des \emph{trasactions conflictuelles}.
%O The notion of conflict is application-defined and transitive, forming an equivalence relation.
La notion de conflit est définie par l'application et est transitive, formant une relation d'équivalence.
%O In our cryptocurrency application, transactions that spend the same funds (\emph{double-spends}) conflict, and form a \emph{conflict set}

%(Figure~\tronly{\ref{fig:dag-conflict-set}}{\ref{fig:dag-cd}})
(shaded regions in Figure~\ref{fig:dag-cd}), out of which only a single one can be accepted.
Note that the conflict set of a virtuous transaction is always a singleton.
%Note that the graph structure of transactions that spend depend on each other, also known as the UTXO graph, is completely independent of the DAG that Avalanche maintains. As a result, any two vertices may be in conflict. Figure~\ref{fig:dag-conflicting-set} shows an example.

### 2020-11-07 paper_fr.tex
S5 - début DAG, création de avalanche_fr.tex

%O To demonstrate the potential of this protocol family, we illustrate a practical
%O peer-to-peer payment system, Avalanche. In effect, Avalanche executes multiple Snowball (one from the Snow family) instances with the aid of a Directed Acyclic Graph (DAG). The DAG serves to piggyback multiple instances, reducing the cost from $\Oh{\log{n}}$ to $\Oh{1}$ per node and streamlining the path where there are no conflicting transactions.
Pour démontrer le potentiel de cette famille de protocoles, nous l'illustrons par un système pratique de paiement pair à pair, Avalanche. Dans les faits, Avalanche exécute plusieurs instances de Snowball (l'un des protocoles de la famille Snow) avec l'aide d'un graphe orienté acyclique (\emph{Directed Acyclic Graph} ou DAG). Le DAG sert à empiler plusieurs instances en réduisant le coût de $\Oh{\log{n}}$ à $\Oh{1}$ par nœud et en optimisant le chemin où il n'existe pas de transactions contradictoires.
%This combination of the best features of traditional and Nakamoto consensus involves one significant tradeoff: liveness for \editchange{equivocating proposals}. %conflicting transactions.
%
%\editchange{Specifically, the consensus protocol from the Snowball family does not guarantee liveness in all cases. It does, however, guarantee logarithmic expected rounds for termination given a propoer initialization heuristic (Section~\ref{sec:liveness}). It also guarantees fast termination when the proposed value is unanimous. As the core part of Avalanche, it guarantees liveness only for \emph{virtuous} transactions, i.e.\ those issued by correct clients and thus guaranteed not to conflict with other transactions.}

@@

%O     \item \textbf{P1. Safety.} When decisions are made by any two correct nodes, they decide on conflicting transactions with negligible probability ($\leq \varepsilon$).
     \item \textbf{P1. Sûreté.} Quand les décisions sont prises par deux nœuds corrects, ils décident des transactions contradictoires avec une probabilité négligeable ($\leq \varepsilon$).
    % \item \textbf{P2. Liveness.} Transactions will be accepted by every correct node in $\Oh{\log n}$ rounds if either the observed Byzantine presence $f'$, where $f' < f$, is sufficiently small, or the network is initialized in a univalent state.
%O     \item \textbf{P2. Liveness (Upper Bound).} Snow protocols terminate with a strictly positive probability within $t_{max}$ rounds.  
     \item \textbf{P2. Vitalité (limite supérieure).} Les protocoles Snow se terminent avec une probabilité strictement positive en au plus $t_{max}$ tours.
%O     \item \textbf{P3. Liveness (Strong Form).} If $f \leq \Oh{\sqrt{n}}$, then the Snow protocols terminate with high probability ($\geq 1 - \varepsilon$) in $\Oh{\log{n}}$ rounds. 
     \item \textbf{P3. Vitalité (forme forte).} Si $f \leq \Oh{\sqrt{n}}$, alors les protocoles Snow se terminent avec une forte probabilité ($\geq 1 - \varepsilon$) en $\Oh{\log{n}}$ tours.
\end{compactitem}
% Instead of unconditional agreement, our approach guarantees that safety will be upheld with probability $1-\varepsilon$, where the choice of the security parameter $\varepsilon$ is under the control of the system designer and applications.

@@

Using Snowball consensus, we have implemented a bare-bones payment system, Avalanche, which supports Bitcoin transactions. In this section, we describe the design and sketch how the implementation can support the value transfer primitive at the center of cryptocurrencies.
Deploying a full cryptocurrency involves bootstrapping, minting, staking, unstaking,
and inflation control. While we have solutions for these issues, their full
discussion is beyond the scope of this paper, whose focus is centered on the
novel Snow consensus protocol family.
%In this section,
%we focus on how Avalanche can support the value transfer primitive at the center of cryptocurrencies.
%\paragraph{XXX: the original implementation paragraph in Model}

%The Snow family is presented in the form of single-decree, binary consensus for its own theoretical value and simplicity.
%On the other hand, we show later in Section~\ref{sec:implementation} our practical payment system, Avalanche, utilizes multiple Snowball instances to support standard Bitcoin-like transactions. %, by humbly cheating in two ways.
%First, instead of a single replicated state machine (RSM) model, where the system determines a sequence of totally-ordered transactions $T_0, T_1, T_2, \ldots$ issued by any client, we adopt a \emph{parallel consensus model}, where each client interacts independently with its own RSMs and optionally transfers ownership of its RSM to another client. The system establishes only a partial order between dependent transactions.
%Second,
In a cryptocurrency setting, cryptographic signatures enforce that only a key owner is able to create a transaction that spends a particular coin. Since correct clients follow the protocol as prescribed and never double spend coins, in Avalanche, they are guaranteed both safety and liveness for their \emph{virtuous} transactions. In contrast, liveness is not guaranteed for \emph{rogue} transactions, submitted by Byzantine clients, which conflict with one another. Such decisions may stall in the network, but have no safety impact on virtuous transactions.
We show that this is a sensible tradeoff, and that the resulting system is sufficient for building complex payment systems.

\input{avalanche}

>>>>

%O To demonstrate the potential of this protocol family, we illustrate a practical
%O peer-to-peer payment system, Avalanche. In effect, Avalanche executes multiple Snowball (one from the Snow family) instances with the aid of a Directed Acyclic Graph (DAG). The DAG serves to piggyback multiple instances, reducing the cost from $\Oh{\log{n}}$ to $\Oh{1}$ per node and streamlining the path where there are no conflicting transactions.
Pour démontrer le potentiel de cette famille de protocoles, nous l'illustrons par un système pratique de paiement pair à pair, Avalanche. Dans les faits, Avalanche exécute plusieurs instances de Snowball (l'un des protocoles de la famille Snow) avec l'aide d'un graphe orienté acyclique (\emph{Directed Acyclic Graph} ou DAG). Le DAG sert à empiler plusieurs instances en réduisant le coût de $\Oh{\log{n}}$ à $\Oh{1}$ par nœud et en optimisant le chemin où il n'existe pas de transactions conflictuelles.
%This combination of the best features of traditional and Nakamoto consensus involves one significant tradeoff: liveness for \editchange{equivocating proposals}. %conflicting transactions.
%
%\editchange{Specifically, the consensus protocol from the Snowball family does not guarantee liveness in all cases. It does, however, guarantee logarithmic expected rounds for termination given a propoer initialization heuristic (Section~\ref{sec:liveness}). It also guarantees fast termination when the proposed value is unanimous. As the core part of Avalanche, it guarantees liveness only for \emph{virtuous} transactions, i.e.\ those issued by correct clients and thus guaranteed not to conflict with other transactions.}

@@

%O     \item \textbf{P1. Safety.} When decisions are made by any two correct nodes, they decide on conflicting transactions with negligible probability ($\leq \varepsilon$).
     \item \textbf{P1. Sûreté.} Quand les décisions sont prises par deux nœuds corrects, ils décident de transactions conflictuelles avec une probabilité négligeable ($\leq \varepsilon$).
    % \item \textbf{P2. Liveness.} Transactions will be accepted by every correct node in $\Oh{\log n}$ rounds if either the observed Byzantine presence $f'$, where $f' < f$, is sufficiently small, or the network is initialized in a univalent state.
%O     \item \textbf{P2. Liveness (Upper Bound).} Snow protocols terminate with a strictly positive probability within $t_{max}$ rounds.  
     \item \textbf{P2. Vitalité (limite supérieure).} Les protocoles Snow se terminent avec une probabilité strictement positive en au plus $t_{max}$ tours.
%O     \item \textbf{P3. Liveness (Strong Form).} If $f \leq \Oh{\sqrt{n}}$, then the Snow protocols terminate with high probability ($\geq 1 - \varepsilon$) in $\Oh{\log{n}}$ rounds. 
     \item \textbf{P3. Vitalité (forme forte).} Si $f \leq \Oh{\sqrt{n}}$, alors les protocoles Snow se terminent avec une forte probabilité ($\geq 1 - \varepsilon$) en $\Oh{\log{n}}$ tours.
\end{compactitem}
% Instead of unconditional agreement, our approach guarantees that safety will be upheld with probability $1-\varepsilon$, where the choice of the security parameter $\varepsilon$ is under the control of the system designer and applications.

@@

%O Using Snowball consensus, we have implemented a bare-bones payment system, Avalanche, which supports Bitcoin transactions. In this section, we describe the design and sketch how the implementation can support the value transfer primitive at the center of cryptocurrencies.
Grâce au consensus Snowball, nous avons implémenté un système de paiement simpliste, Avalanche, qui supporte les transactions Bitcoin. Dans cette section, nous en décrivons la conception et schématison la manière donc l'implémentation peut supporter la primitive de transfert de valeur au centre des cryptomonnaies.
%O Deploying a full cryptocurrency involves bootstrapping, minting, staking, unstaking,
%O and inflation control. While we have solutions for these issues, their full
%O discussion is beyond the scope of this paper, whose focus is centered on the
%O novel Snow consensus protocol family.
Le déploiement complet d'une cryptomonnaie implique l'initialisation, l'émission, le dépôt, le retrait et le contrôle de l'inflation. Nous disposons de solutions à ces problèmes mais leur description détaillée sort du cadre de cet article, dont l'objet est la nouvelle famille de consensus Snow. % HOULA staking unstaking
%In this section,
%we focus on how Avalanche can support the value transfer primitive at the center of cryptocurrencies.
%\paragraph{XXX: the original implementation paragraph in Model}

%The Snow family is presented in the form of single-decree, binary consensus for its own theoretical value and simplicity.
%On the other hand, we show later in Section~\ref{sec:implementation} our practical payment system, Avalanche, utilizes multiple Snowball instances to support standard Bitcoin-like transactions. %, by humbly cheating in two ways.
%First, instead of a single replicated state machine (RSM) model, where the system determines a sequence of totally-ordered transactions $T_0, T_1, T_2, \ldots$ issued by any client, we adopt a \emph{parallel consensus model}, where each client interacts independently with its own RSMs and optionally transfers ownership of its RSM to another client. The system establishes only a partial order between dependent transactions.
%Second,
%O In a cryptocurrency setting, cryptographic signatures enforce that only a key owner is able to create a transaction that spends a particular coin. Since correct clients follow the protocol as prescribed and never double spend coins, in Avalanche, they are guaranteed both safety and liveness for their \emph{virtuous} transactions. In contrast, liveness is not guaranteed for \emph{rogue} transactions, submitted by Byzantine clients, which conflict with one another. Such decisions may stall in the network, but have no safety impact on virtuous transactions.
Dans une configuration de cryptomonnaie, les signatures cryptographiques imposent que seul le possesseur de la clef soit capable de créer une transaction dépensant un dépôt donné. Comme les clients corrects suivent le protocole tel qu'il est prescrit et ne peuvent pratiquer la double dépense, ils ont la garantie tant de sûreté de fonctionnement que de vitalité pour leurs transactions \emph{vertueuses}. Par contraste, la vitalité n'est pas garantie pour les transactions \emph{malveillantes} soumises par des clients byzantins, qui entrent mutuellement en conflit. Ces décisions peuvent arrêter le réseau mais n'ont pas d'impact sur les transactions vertueuses.
%O We show that this is a sensible tradeoff, and that the resulting system is sufficient for building complex payment systems.
Nous montrons que ce compromis est sensé et que le système qui en résulte suffit à bâtir des systèmes de paiement complexes.

\input{avalanche_fr}
